Experiences designing and building a PRAGMA Cloud Scheduler

### 2015-08-13 paper.tex
Revised intro slightly and added requirements

\maketitle
\begin{abstract}
The Pacific Rim Application and Grid Middleware Assembly (PRAGMA) is a community of individuals and sites from around the Pacific Rim that actively collaborate to enable scientific expeditions in areas like biodiversity and lake ecology.  Over the past X years, the technology focus for PRAGMA partners has shifted to cloud and software defined networking as enabling technologies.  During PRAGMA 27, the Resources Working group discussed rebooting the persistent PRAGMA Testbed as a way for sites to contribute and use shared resources, leveraging technologies such as PRAGMA Boot, Personal Cloud Controller (PCC), and overlay networks.  To make PRAGMA resource sharing easier, a lightweight scheduler was proposed and discussed as a way to enable access to the resources and to manage resource reservations.  This short paper discusses the design process and initial prototype of a simple cloud scheduler for PRAGMA.

@@

The first phase of the PRAGMA cloud testbed started with three sites and focused around the creation of application specific virtual machines and making them portable to  different cloud hosting environments like Rocks~\cite{rocks}, OpenNebula~\cite{opennebula}, and Eucalyptus~\cite{eucalyptus}.
One early example was an AutoDoc~\cite{autodock} virtual machine created for the Avian Flu research team in 2011~\cite{pragmaReport2011}.   By 2012, the cloud testbed had ten sites with a total of 367 CPUs, 2.5 terabytes of memory, and 657 terabytes of online storage.  To make it easier for users to assemble a multi-node virtual environment for running their scientific experiments, PRAGMA shifted its  focus to the creation and management of virtual clusters in 2013.  The following technologies have been explored to facilitate both the operation and use of virtual clusters on the PRAGMA cloud testbed.

@@

In late 2014 during the PRAGMA27 workshop, a lightweight scheduler was proposed to coordinate access to the above technologies.  The next section discusses the general requirements of the cloud scheduler and Section~\ref{Sec:Design} discusses the design options that were considered and why a simple calendar solution was selected.  A summary of calendar systems that were considered is provided in Section~\ref{Sec:Calendars}.  Section~\ref{Sec:Pilot} discusses our early implementation with one of the calendar systems called Booked and Section~\ref{Sec:Conclusions} concludes with our planned future work.

@@

Sites willing to participate will have to install a PRAGMA Package - a very small set of components.
Users can request access to resources similar to requesting a room (i.e., a room reservation system)
PCC would be used to automatically start and monitor user's VMs slash VC via pragma\_boot up at sites 
Only support 10s of users

@@

@@

%ACKNOWLEDGMENTS are optional
%\section{Acknowledgments}

>>>>

\maketitle
\begin{abstract}
The Pacific Rim Application and Grid Middleware Assembly (PRAGMA) is a community of individuals and sites from around the Pacific Rim that actively collaborate to enable scientific expeditions in areas like biodiversity and lake ecology.  Over the past four years, the technology focus for PRAGMA partners has shifted to cloud and software defined networking as enabling technologies.  During PRAGMA 27, the Resources Working group discussed rebooting the persistent PRAGMA Testbed as a way for sites to contribute and use shared resources, leveraging technologies such as PRAGMA Boot, Personal Cloud Controller (PCC), and overlay networks.  To make PRAGMA resource sharing easier, a lightweight scheduler was proposed and discussed as a way to enable access to the resources and to manage resource reservations.  This short paper discusses the design process and initial prototype of a simple cloud scheduler for PRAGMA.

@@

The first phase of the PRAGMA cloud testbed started with three sites and focused around the creation of application specific virtual machines and making them portable to  different cloud hosting environments like Rocks~\cite{rocks}, OpenNebula~\cite{opennebula}, and Eucalyptus~\cite{eucalyptus}.
One early example was an AutoDoc~\cite{autodock} virtual machine created for the Avian Flu research team in 2011~\cite{pragmaReport2011}.   By 2012, the cloud testbed had ten sites with a total of 367 CPUs, 2.5 terabytes of memory, and 657 terabytes of online storage.  In addition to the previously managed cloud tools, To make it easier for users to assemble a multi-node virtual environment for running their scientific experiments, PRAGMA shifted its  focus to the creation and management of virtual clusters in 2013.  The following technologies have been explored to facilitate both the operation and use of virtual clusters on the PRAGMA cloud testbed.

@@

In late 2014 during the PRAGMA27 workshop, a lightweight scheduler was proposed to coordinate running on the different cloud deployments and leverage the above technologies.  The next section discusses the general requirements of the cloud scheduler and Section~\ref{Sec:Design} discusses the design options that were considered and why a simple calendar solution was selected.  A summary of calendar systems that were considered is provided in Section~\ref{Sec:Calendars}.  Section~\ref{Sec:Pilot} discusses our early implementation with one of the calendar systems called Booked and Section~\ref{Sec:Conclusions} concludes with our planned future work.

@@

PRAGMA had the following three key requirements for a cloud testbed scheduler:

\textbf{Low participation overhead}:  One key requirement for the cloud testbed scheduler was that it be lightweight and require  minimal effort and minimal expertise for a site to add their cloud deployment to the list of schedulable resources.   Similarly the scheduler would need to work across multiple cloud deployment tools (e.g., Rocks, OpenNebula, and OpenStack~\cite{openstack}) so that sites do not have to learn and install one specific cloud deployment tool.  Sites  just need to run a SSH service and install the pragma\_boot package, which is a small Python package and can be installed via RPM and can be configured in less than an hour.

\textbf{Easy to use}:  To execute a virtual cluster on the PRAGMA cloud testbed currently, users needs to contact each site individually to gain access to their cloud deployment, upload their images, and if pragma\_boot is not available, perform all the steps manually to launch their virtual cluster.   The design of the PRAGMA cloud scheduler should first simplify that process by ensuring pragma\_boot gets deployed more broadly and ported it to more cloud systems as needed.   The scheduler should also provide a simple Web interface for users to see the available resources at sites, construct their virtual cluster, and manage their images.

\textbf{Scale to tens of users:}  While the PRAGMA community was composed of nineteen active institution sites in 2014, the number of expected users for the cloud testbed is in the tens of users, not hundreds nor thousands.  Therefore, our scheduler can be simplified and we can give high priority to the requirements of low participation overhead and easy of use.

Users can request access to resources similar to requesting a room (i.e., a room reservation system)
PCC would be used to automatically start and monitor user's VMs slash VC via pragma\_boot up at sites 
Only support 10s of users

@@

ENT
one controller at participating site
way to slice switch? specialized controller? 
AutoVFlow by Hiroaki Yamanaka (NICT)
For users interested in running VMs, cloud middleware (OpenStack, CloudStack, etc) should take care of the slicing.
 Recurring problem: users tend to forget and leave idle VMs running.
For users interested in cloud middleware research/development, and/or access to bare-metal hardware (e.g., ENT), scheduling of resources and coordination among users will be needed.
Network slicing technologies (Auto-VFlow, FlowVisor, FlowSpace Firewall, etc) when deployed could help
Use PBS as a scheduler (or Condor or Slurm?)
Nimbus have a module that interfaces with PBS
 Use PBS to schedule interactive sessions on bare metal
Use a shared calendar so that users can make resource reservations (`a la Grid?5000)
ORCA from GENI/Planetlab? 
Leases

@@

%ACKNOWLEDGMENTS are optional
\section{Acknowledgments}

Thank Mauricio and Jose.
